---
title: "How to Create an Interactive WebGL Network Graph Using R"
author: "Max Woolf (@minimaxir)"
date: "November 15th, 2016"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: yes
---

This R Notebook is the complement to my blog post [What Percent of the Top-Voted Comments in Reddit Threads Were Also 1st Comment?](http://minimaxir.com/2016/11/first-comment/).

This notebook is licensed under the MIT License. If you use the code or data visualization designs contained within this notebook, it would be greatly appreciated if proper attribution is given back to this notebook and/or myself. Thanks! :)

# Setup

Setup the R packages.

```{r}
library(dplyr)
library(nycflights13)
library(igraph)
library(intergraph)
library(sna)
library(ggplot2)
library(ggnetwork)
library(viridis)
library(plotly)
library(htmlwidgets)
```

The `nycflights13` package contains a `flights` dataset.

```{r}
flights %>% head()
```

There are **`r flights %>% nrow() %>% format(big.mark=",")`** flights in the dataset.

# Build the Network

Getting the edge weights is a `dplyr` aggregation.

```{r}
df_edges <- flights %>% group_by(origin, dest) %>% summarize(weight = n())
df_edges %>% arrange(desc(weight)) %>% head()
```

There are **`r df_edges %>% nrow() %>% format(big.mark=",")`** total edges.

Add a colors column to edge for each `origin` which will eventually be used for final ggplot.

```{r}
# blue, red, green
colors = c("#3498db", "#e74c3c", "#2ecc71")

# seting alphabetical order; allows for predictable ordering later
origins = c("EWR", "JFK", "LGA")

df_colors = tbl_df(data.frame(origin=origins, color=origins))
df_edges <- df_edges %>% left_join(df_colors)

df_edges %>% arrange(desc(weight)) %>% head()
```


```{r}
net <- graph.data.frame(df_edges, directed = T)
V(net)$degree <- centralization.degree(net)$res
V(net)$weighted_degree <- graph.strength(net)
V(net)$color_v <- c(origins, rep("Others", gorder(net) - length(colors)))

net
```

Write specialized hovertext for each vertex. Note that airport attributes must be mapped to same order as vertices.


```{r}

df_airports <- data.frame(vname=V(net)$name) %>% left_join(airports, by=c("vname" = "faa"))

V(net)$text <- paste(V(net)$name,
                       df_airports$name,
                       paste(format(V(net)$weighted_degree, big.mark=",", trim=T), "Flights"),
                        sep = "<br>")

V(net)$text %>% head()
```

Add latitudes/longitudes to both vertices and edges for spatial map;

```{r}
V(net)$lat <- df_airports$lat
V(net)$lon <- df_airports$lon

# gives to/from locations; map to corresponding ending lat/long
end_loc <- data.frame(ename=get.edgelist(net)[,2]) %>% left_join(airports, by=c("ename" = "faa"))

E(net)$endlat <- end_loc$lat
E(net)$endlon <- end_loc$lon
```

# Plotting the Network Graph

Use `ggnetwork` to transform the network to a `ggplot` friendly format.

```{r}
df_net <- ggnetwork(net, layout = "fruchtermanreingold", weights="weight", niter=5000, arrow.gap=0)
df_net %>% head()
```

```{r}
plot <- ggplot(df_net, aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = color), size=0.4, alpha=0.25) +
    geom_nodes(aes(color = color_v, size = degree, text=text)) +
    ggtitle("Network Graph of U.S. Flights Outbound from NYC in 2013") +
    scale_color_manual(labels=c("LGA", "EWR", "JFK", "Others"), 
                         values=c(colors, "#1a1a1a"), name="Airports") +
    guides(size=FALSE) +
    theme_blank()

plot
```

```{r}
plot %>% ggplotly(tooltip="text") %>% toWebGL()
```

Save interactive plot locally to disk using `htmlwidgets` for uploading to my website. (this only saves the data/layout; you will need to provide the relevant plot.ly javascript on your own website)

```{r}
plot %>% ggplotly(tooltip="text", width="100%", height=600) %>%
    saveWidget("ggplot-graph-1.html", selfcontained=F, libdir="plotly")
```

## Plot by Physical Location

```{r}
plot <- ggplot(df_net, aes(x = lon, y = lat, xend = endlon, yend = endlat)) +
    geom_edges(aes(color = color), size=0.4, alpha=0.25) +
    geom_nodes(aes(color = color_v, size = degree, text=text)) +
    ggtitle("Locations of U.S. Flights Outbound from NYC in 2013") +
    scale_color_manual(labels=c("LGA", "EWR", "JFK", "Others"), 
                         values=c(colors, "#1a1a1a"), name="Airports") +
    guides(size=FALSE) +
    theme_blank()

plot
```

```{r}
plot %>% ggplotly(tooltip="text") %>% toWebGL()
```

```{r}
plot %>% ggplotly(tooltip="text", width="100%", height=600) %>%
    saveWidget("ggplot-graph-2.html", selfcontained=F, libdir="plotly")
```

# LICENSE

The MIT License (MIT)

Copyright (c) 2016 Max Woolf

Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the "Software"), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.