---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(nycflights13)
library(igraph)
library(intergraph)
library(sna)
library(ggplot2)
library(ggnetwork)
library(viridis)
library(plotly)
```

```{r}
flights %>% head()
```

There are `r flights %>% nrow() %>% format(big.mark=",")` flights.

```{r}
df_edges <- flights %>% group_by(origin, dest) %>% summarize(weight = n())
df_edges %>% arrange(desc(weight)) %>% head()
```

There are `r df_edges %>% nrow() %>% format(big.mark=",")` total edges.

Add colors to edge for each `origin` which will be used for final ggplot.

```{r}
# blue, red, green
colors = c("#3498db", "#e74c3c", "#2ecc71")

# seting alphabetical order; allows for predictable ordering later
origins = c("EWR", "JFK", "LGA")

df_colors = tbl_df(data.frame(origin=origins, color=origins))
df_edges <- df_edges %>% left_join(df_colors)

df_edges %>% arrange(desc(weight)) %>% head()
```


```{r}
net <- graph.data.frame(df_edges, directed = T)
V(net)$degree <- centralization.degree(net)$res
V(net)$weighted_degree <- graph.strength(net)
V(net)$color_v <- c(origins, rep("Others", gorder(net) - length(colors)))

net
```

Write specialized hovertext for each vertex. Note that airport attributes must be mapped to same order as vertices.


```{r}

df_airports <- data.frame(vname=V(net)$name) %>% left_join(airports, by=c("vname" = "faa"))

V(net)$text <- paste(V(net)$name,
                       df_airports$name,
                       paste(format(V(net)$weighted_degree, big.mark=",", trim=T), "Flights"),
                        sep = "<br>")

V(net)$text %>% head()
```

Add latitudes/longitudes to both vertices and edges for spatial map;

```{r}
V(net)$lat <- df_airports$lat
V(net)$lon <- df_airports$lon

# gives to/from locations; map to corresponding ending lat/long
end_loc <- data.frame(ename=get.edgelist(net)[,2]) %>% left_join(airports, by=c("ename" = "faa"))

E(net)$endlat <- end_loc$lat
E(net)$endlon <- end_loc$lon
```

# Plotting the Network Graph

```{r}
df_net <- ggnetwork(net, layout = "fruchtermanreingold", weights="weight", niter=5000, arrow.gap=0)
df_net %>% head()
```

```{r}
plot <- ggplot(df_net, aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = color), size=0.4, alpha=0.25) +
    geom_nodes(aes(color = color_v, size = degree, text=text)) +
    ggtitle("Network Graph of U.S. Flights Outbound from NYC in 2013") +
    scale_color_manual(labels=c("LGA", "EWR", "JFK", "Others"), 
                         values=c(colors, "#1a1a1a"), name="Airports") +
    guides(size=FALSE) +
    theme_blank()

plot
```

```{r}
plot %>% ggplotly(tooltip="text") %>% toWebGL()
```

## Plot by Physical Location

```{r}
plot <- ggplot(df_net, aes(x = lon, y = lat, xend = endlon, yend = endlat)) +
    geom_edges(aes(color = color), size=0.4, alpha=0.25) +
    geom_nodes(aes(color = color_v, size = degree, text=text)) +
    ggtitle("Locations of U.S. Flights Outbound from NYC in 2013") +
    scale_color_manual(labels=c("LGA", "EWR", "JFK", "Others"), 
                         values=c(colors, "#1a1a1a"), name="Airports") +
    guides(size=FALSE) +
    theme_blank()

plot
```

```{r}
plot %>% ggplotly(tooltip="text") %>% toWebGL()
```
